---
title: "Data Cleaning Exploration"
author: "Phoebe Abramowitz"
date: "3/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Upload Files:
```{r}
setwd("~/Desktop/Berkeley/Spring19/stat_154/154project")
source("./data_cleaning .R")
```

Network retrieved data until the end but sometimes the connection failed, logger has data from when the connection failed but not always until the end

```{r}
#Ask how to read this, or manually change the format into something I can read with delim then cut 
#datetime <- read.delim("./data/sonoma-dates", header=FALSE)
```
Temperature:
```{r}
#why are min & max NA?
range(data_net$humid_temp)
range(data_log$humid_temp)
summary(data_net$humid_temp)
summary(data_log$humid_temp)
ggplot(data_net, aes(x=humid_temp)) + geom_histogram()
ggplot(data_log, aes(x=humid_temp)) + geom_histogram()
```
Relative Humidity:
```{r}
range(data_net$humidity)
range(data_log$humidity)
summary(data_net$humidity)
#Same Scale just with crazy lower outlier
summary(data_log$humidity)
ggplot(data_net, aes(x=humidity)) + geom_histogram()
ggplot(data_log, aes(x=humidity)) + geom_histogram()
```
Incident PAR:hamatop? wavelength from 350 to 700 nm
```{r}
#range for net is in the 90s, for log in the 40s
range(data_net$hamatop)
range(data_log$hamatop)
summary(data_net$hamatop)
summary(data_log$hamatop)
ggplot(data_net, aes(x=hamatop)) + geom_histogram()
ggplot(data_log, aes(x=hamatop)) + geom_histogram()
```
Reflective PAR:check column use here
```{r}
range(data_net$hamabot)
range(data_log$hamabot)
summary(data_net$hamabot)
#similar in scale except for Max. >1/2 zeros bc median=0
summary(data_log$hamabot)
ggplot(data_net, aes(x=hamabot)) + geom_histogram()
ggplot(data_log, aes(x=hamabot)) + geom_histogram()
```
Voltage-measure of Battery:
```{r}
range(data_net$voltage)-range(data_log$voltage)
#range of voltage is very inconsistant
#voltage is clearly different scales between log and net
#loig data matches the paper
```
epoch, nodeid, result_time-Distinguishing between points
```{r}
#range of epoch is very inconsistant
range(data_net$epoch)-range(data_log$epoch)

#Why does nodeid go so high in log? 
range(data_net$nodeid)-range(data_log$nodeid)
```
Choose to combine net and log, or find a way to use all? Either can work to become the "Main Table"
What is epoch?
Are all the missing Values NA? How to get range to stop giving NA
When should zero values be NAs?
How to use the HUmid adjustment column?

