---
title: "Data Exploration"
author: "Phoebe Abramowitz"
date: "3/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
```
Make some pairwise scatterplots of some variables. Pick a reasonable time period.
Explain your choice and describe your findings.
```{r}
#good head 
head(data_main)

# Reasonable time period is within range of 1000<epoch<3500, because most values occur there, not first
#99,000 rows(too big?)
data_sub <- filter(filter(data_main,epoch<3500),(epoch>1000))
nrow(data_sub)

#I'm using epoch--the sample number-- as a substitute for time & date throughout, becuase it's numeric

#The number of the nodeid is not meaningfully correlated to height
ggplot(data_sub)+
  geom_point(aes(x=height, y=nodeid, alpha=0.3),color="darkgoldenrod")+
  ggtitle("nodeid by height")+
  theme_minimal()

# Humidity
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=humid, alpha=0.3),color="firebrick2")+
  ggtitle("Humidity Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=humid, alpha=0.3),color="violet")+
  ggtitle("Humidity by Height")+
  theme_minimal()

# Temperature

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=temp, alpha=0.2),color="orangered1")+
  ggtitle("Temperature Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=temp, alpha=0.3),color="deeppink1")+
  ggtitle("Temperature by Height")+
  theme_minimal()

#Reflective PAR
#loosely indicates cycles of daylight
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=reflect_PAR, alpha=0.3),color="deeppink1")+
  ggtitle("Reflective PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=reflect_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Reflective PAR by Height")+
  theme_minimal()

# Direct PAR
```
b)Are any of the predictors associated with Incident PAR? If so, explain the relationship

Incident PAR is visibly cyclical over time. The relationship can be explained by the daily daylight cycle of the sun rising and setting, since Incident PAR measures direct sunlight. 
```{r}
#Clearly shows daylight cycle
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=incident_PAR, alpha=0.3),color="orangered1")+
  ggtitle("Incident PAR Over Time")+
  theme_minimal()

#These two plots show that certain nodes are exposed to more direct Sunlight, but not necessarily by height
ggplot(data_sub)+
  geom_point(aes(x=height, y=incident_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Incident PAR by Height")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=nodeid, y=incident_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Incident PAR by Node")+
  theme_minimal()

#Incident & Reflective PAR(interesting how little these seems correlated)
ggplot(data_sub)+
  geom_point(aes(x=incident_PAR, y=reflect_PAR, alpha=0.3),color="darkgoldenrod")+
  ggtitle("Incident vs Refletive PAR")+
  theme_minimal()

```

Consider each variable as a time series and look at its temporal trend. Generate such
plots (value vs time) with height as color cue for at least four variables (Temperature,
Relative Humidity, Incident PAR and Reflected PAR). You can do it for different time
scales (during an hour, during a day or during the entire experiment). However, at
least the plots with days as x-axis are required. Comment on the range, continuity and
strange behaviors in these variables.
```{r}
#Is it okay to use epoch?
#Is result_time in a format that r can recognize as chronological? In plot below, no
ggplot(data_sub)+
  geom_point(aes(x=result_time, y=epoch, alpha=0.3))+
  ggtitle("epoch by result time")+
  theme_minimal()

#Time Series Plots
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=incident_PAR, alpha=0.3,color=height))+
  ggtitle("Incident PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=reflect_PAR, alpha=0.3,color=height))+
  ggtitle("Reflective PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=temp, alpha=0.3,color=height))+
  ggtitle("Temperature Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=humid, alpha=0.3,color=height))+
  ggtitle("Humidity Over Time")+
  theme_minimal()
```
```{r}
ggplot(data_main %>% filter(result_time <= as.Date("2004-11-10"))) +
  geom_point(aes(x=result_time, y=epoch))
```

```{r}
# PCA analysis
dat <- data_sub %>%
  select(c("epoch", "temp", "humid", "incident_PAR", "reflect_PAR", "height"))

pca <- prcomp(na.omit(dat), scale. = TRUE)
summary(pca)
```

```{r}
loadings <- pca$rotation
scores <- pca$x
eigenvalues <- pca$sdev^2
eigenvalues
sum(eigenvalues)

loadings[1:5,]
scores[1:5,]
```
```{r}
eigs_cum = cumsum(eigenvalues) / sum(eigenvalues)
ggplot() + geom_point(aes(x = 1:length(eigenvalues), y=eigs_cum)) +
  labs(x = "first PCs", y = "fraction of total variance explained") +
  ggtitle("Screeplot") + theme_minimal()
```
```{r}
biplot(x = scores[1:1000 ,1:2], y = loadings)
```

```{r}
# K-means
K5 <- kmeans(na.omit(dat), 5)
table(K5$cluster)
```

