---
title: "Data Exploration"
author: "Phoebe Abramowitz"
date: "3/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
```
Make some pairwise scatterplots of some variables. Pick a reasonable time period.
Explain your choice and describe your findings.
```{r}
data_main %>% 
  filter(result_time >= as.Date("2004-05-23") & result_time <= as.Date("2004-05-30"))

# using this subsection of the data, we get an accurate reading of result_time, by replacing data_log result time with respective data_net result_time using the epoch as an identifier (we noticed epoch and result time are linearly correlated). The last week of May experiences higher levels of daylight hours as the seasons begin to shift and the summer solstice gets closer. We postulate this extended day light hours will have certain affects on the 
```

```{r}
#good head 
head(data_main)

# Reasonable time period is within range of 1000<epoch<3500, because most values occur there, not first
#99,000 rows(too big?)
data_sub <- data_main %>% 
  filter(result_time >= as.Date("2004-05-23") & result_time <= as.Date("2004-05-30"))
nrow(data_sub)

#I'm using epoch--the sample number-- as a substitute for time & date throughout, becuase it's numeric

#The number of the nodeid is not meaningfully correlated to height
ggplot(data_sub)+
  geom_point(aes(x=height, y=nodeid),color="darkgoldenrod")+
  ggtitle("nodeid by height")+
  theme_minimal()

# Humidity
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=humid, alpha=0.3),color="firebrick2")+
  ggtitle("Humidity Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=humid, alpha=0.3),color="violet")+
  ggtitle("Humidity by Height")+
  theme_minimal()

# Temperature

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=temp, alpha=0.2),color="orangered1")+
  ggtitle("Temperature Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=temp, alpha=0.3),color="deeppink1")+
  ggtitle("Temperature by Height")+
  theme_minimal()

#Reflective PAR
#loosely indicates cycles of daylight
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=reflect_PAR, alpha=0.3),color="deeppink1")+
  ggtitle("Reflective PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=height, y=reflect_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Reflective PAR by Height")+
  theme_minimal()

# Direct PAR
```
b)Are any of the predictors associated with Incident PAR? If so, explain the relationship

Incident PAR is visibly cyclical over time. The relationship can be explained by the daily daylight cycle of the sun rising and setting, since Incident PAR measures direct sunlight. 
```{r}
#Clearly shows daylight cycle
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=incident_PAR, alpha=0.3),color="orangered1")+
  ggtitle("Incident PAR Over Time")+
  theme_minimal()

#These two plots show that certain nodes are exposed to more direct Sunlight, but not necessarily by height
ggplot(data_sub)+
  geom_point(aes(x=height, y=incident_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Incident PAR by Height")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=nodeid, y=incident_PAR, alpha=0.3), color="deeppink1")+
  ggtitle("Incident PAR by Node")+
  theme_minimal()

#Incident & Reflective PAR(interesting how little these seems correlated)
ggplot(data_sub)+
  geom_point(aes(x=incident_PAR, y=reflect_PAR, alpha=0.3),color="darkgoldenrod")+
  ggtitle("Incident vs Refletive PAR")+
  theme_minimal()

```

Consider each variable as a time series and look at its temporal trend. Generate such
plots (value vs time) with height as color cue for at least four variables (Temperature,
Relative Humidity, Incident PAR and Reflected PAR). You can do it for different time
scales (during an hour, during a day or during the entire experiment). However, at
least the plots with days as x-axis are required. Comment on the range, continuity and
strange behaviors in these variables.
```{r}
#Is it okay to use epoch?
#Is result_time in a format that r can recognize as chronological? In plot below, no
ggplot(data_sub)+
  geom_point(aes(x=result_time, y=epoch, alpha=0.3))+
  ggtitle("epoch by result time")+
  theme_minimal()

#Time Series Plots
ggplot(data_sub)+
  geom_point(aes(x=epoch, y=incident_PAR, alpha=0.3,color=height))+
  ggtitle("Incident PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=reflect_PAR, alpha=0.3,color=height))+
  ggtitle("Reflective PAR Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=temp, alpha=0.3,color=height))+
  ggtitle("Temperature Over Time")+
  theme_minimal()

ggplot(data_sub)+
  geom_point(aes(x=epoch, y=humid, alpha=0.3,color=height))+
  ggtitle("Humidity Over Time")+
  theme_minimal()
```

Plot Showing we can use epoch to represent time
```{r}
ggplot(data_main %>% filter(result_time <= as.Date("2004-11-10"))) +
  geom_point(aes(x=result_time, y=epoch))
```

```{r}
# PCA analysis
dat <- data_sub %>%
  select(c("temp", "humid", "incident_PAR", "reflect_PAR", "height"))

pca <- prcomp(na.omit(dat), scale. = TRUE)
summary(pca)
```

```{r}
loadings <- pca$rotation
scores <- pca$x
eigenvalues <- pca$sdev^2
eigenvalues
sum(eigenvalues)

loadings[1:5,]
scores[1:5,]
```
```{r}
eigs_cum = cumsum(eigenvalues) / sum(eigenvalues)
ggplot() + geom_point(aes(x = 1:length(eigenvalues), y=eigs_cum)) +
  labs(x = "first PCs", y = "fraction of total variance explained") +
  ggtitle("Screeplot") + theme_minimal()
# use first two principal components because they both have variances greater than one (Kaiser's criterion)
```

```{r}
# principal component scatter plots
ggplot() + 
  geom_point(aes(x = scores[, 1], y=scores[, 2], alpha=0.1)) +
   labs(x = "PC1", y = "PC2") + ggtitle("PC1 vs. PC2")

ggplot() + 
  geom_point(aes(x = scores[, 1], y=scores[, 3], alpha=0.1)) +
   labs(x = "PC1", y = "PC3") + ggtitle("PC1 vs. PC3")

ggplot() + 
  geom_point(aes(x = scores[, 1], y=scores[, 4], alpha=0.1)) +
   labs(x = "PC1", y = "PC4") + ggtitle("PC1 vs. PC4")

ggplot() + 
  geom_point(aes(x = scores[, 1], y=scores[, 5], alpha=0.1)) +
   labs(x = "PC1", y = "PC5") + ggtitle("PC1 vs. PC5")
```
```{r}
# trying to do kmeans on PC1 and PC2
pca_kmeans <- kmeans(scores[,1:2], 3)
plot(dat, col=(pca_kmeans$cluster+1), main="K-Means Results on First Two Principal Components (K=3)", xlab="", ylab="", pch=20, cex=2)
```

```{r}
biplot(pca, scale=0)
#biplot(x = scores[1:500, 1:2], y=loadings)
```

```{r}
# K-means
K5 <- kmeans(na.omit(dat), 5)
table(K5$cluster)
```

```{r}
# 5a.
ggplot(data_main) +
  geom_histogram(aes(x=incident_PAR),color="darkblue", fill="lightblue") +
  ggtitle("Incident PAR") +
  scale_y_sqrt() +
  theme_minimal()

ggplot(data_main) +
  geom_histogram(aes(x=reflect_PAR),color="darkblue", fill="lightblue") +
  ggtitle("Reflected PAR") +
  scale_y_sqrt() +
  theme_minimal()
```


```{r}
# 5b.
# What messages are figures 3c and 3d trying to convey? Do they convey the right message? If not, generate a new plot with the same data (compare to some plots in figure 4)

# They are 'projecting onto height and value dimensions' to look for spatial trends. 3c shows the distribution of all the readings taken by each sensor at each height. 
# We see a very different spatial trend in the temperature and humidity readings: nothing. Every sensor reached practically every point in the space of possible temperature and humidity readings. This suggests that the amount of variation over time overwhelmed the amount of variation over space. However, instead of just collapsing the temporal variation, we can explicitly remove it and focus more closely on the spatial trends. At every timestep, we take the mean of all the sensor readings. We can subtract the timestep mean from each sensor reading, and examine the distributions of the differences 3d 

ggplot(data_main) +
  geom_point(aes(x=temp, y=height, alpha=0.1)) +
  theme_minimal()

```

